Algebraic File Synchronization: Adequacy and Completeness

### 2017-10-06 p2/conclusion.tex
Remove unused symbols; clarify text

And finally, we hope that this work, together with \cite{NREC}, provides
a blueprint of constructing an algebra of commands for different storage protocols
(e.g. XML trees, mailbox folders, generic relational databases, etc.),
and of demonstrating the adequacy and completeness of update and conflict detection and reconciliation
algorithms defined over it.
This, in turn, can offer formal verification of the algorithms underlying
specific implementations in a variety of synchronizers.
Alternatively, by generalizing the parent--child relationships between filesystem nodes,
the demonstrated properties of minimal sequences of commands
and conditional operation of sequences ($\workssign$)
may also contribute to future research into algebraic structures
constrained by predefined sparse connections between their elements.

>>>>

And finally, we hope that this work, together with \cite{NREC}, provides
a blueprint of constructing an algebra of commands for different storage protocols
(e.g. XML trees, mailbox folders, generic relational databases, etc.),
and of demonstrating the adequacy and completeness of update and conflict detection and reconciliation
algorithms defined over it.
This, in turn, can offer formal verification of the algorithms underlying
specific implementations in a variety of synchronizers.
Alternatively, by generalizing the parent--child relationships between filesystem nodes,
the demonstrated properties of minimal sequences of commands
and conditional operation of sequences ($\worksmeqsign$) % TODO
may also contribute to future research into algebraic structures
constrained by predefined sparse connections between their elements.

### 2017-10-06 p2/ordering.tex
New script to enumerate references
Remove unused symbols; clarify text

\begin{mylem}\label{simple_distant_pairs}
For any simple sequence $S\nequiv\cbrk$
if $\cxynv\cc\czwmv$ is a subsequence of $S$ and $n=\parent(m)$ or $\parent(n)=m$,
then $\cxynv\cc\czwmv\nequiv\cbrk$.
\end{mylem}
\begin{proof}
Informally speaking, the proposition means that
the two commands must form a construction or destruction pair even
if they are not next to each other.
This is true because in simple sequences
there cannot be another command on $n$ or $m$,
so if the commands are incompatible, no command between them can change that.
In fact, this \namecref{simple_distant_pairs} is the general case
of \cref{ax_directchild_breaks,ax_directparent_breaks}.

% BACK_TO_THE_FS TODO
% TODO Review any proof depending on "nothing changes FS at n until...".
% TODO Move to rules
% TODO APPENDIX
Formally, we prove our proposition by contradiction,
and in this version of the proof we reach back to our filesystem model.
We assume $S\nequiv\cbrk$ and therefore there is a $\FS$ for which $S\aFS\neq\fsbroken$,
but $S$ contains commands on a node and its parent that do not form a construction or destruction pair.
Select two such commands, $\cxynv$ and $\czwmv$,
and split $S$ around them into three parts:
\[ S = S_0 \cc \cxynv \cc S_1 \cc \czwmv \cc S_2. \]
As $S$ is simple, there cannot be any commands on $n$ or $m$ in $S_1$,
and therefore $(S_0\cc\cxynv\cc S_1)\FS(n) = y$,
and $[(S_0\cc\cxynv\cc S_1)\FS(m)] = [S_0\FS(m)]$.
If $[z]\neq[S_0\FS(m)]$, then $S\aFS=\fsbroken$, which is a contradiction.

If $n=\parent(m)$, then 
we know $(S_0\cc\cxynv\cc S_1)\FS(n) = y$, and as $[z]\neq[w]$, either
$[(S_0\cc\cxynv\cc S_1)\FS(m)]=[z]\neq[\empt]$,
or $[(S_0\cc\cxynv\cc S_1\cc\czwmv)\FS(m)]=[w]\neq[\empt]$.
Therefore $y$ must be $\vald$, as otherwise the tree property would be violated
when applying $\czwmv$,
and $[x]\neq[\vald]$ as $\cxynv$ is not an assertion command.
As $[S_0\FS(n)]=[x]\neq[\vald]$, we know
$S_0\FS(m)=\empt$ as otherwise the tree property would be violated.
Therefore $[S_0\FS(m)]=[z]=[\empt]$, which, combined with $y=\vald$
means that $\cxynv\cc\czwmv\nequiv\cbrk$ is a construction pair, which is a contradiction.

If $\parent(n)=m$, then 
as $(S_0\cc\cxynv\cc S_1)\FS(n) = y$, and either
$[(S_0\cc\cxynv\cc S_1)\FS(m)]=[z]\neq[\vald]$
or $[(S_0\cc\cxynv\cc S_1\cc\czwmv)\FS(m)]=[w]\neq[\vald]$,
$y$ must be $\empt$, and $[x]\neq[\empt]$.
As $[S_0\FS(n)]=[x]\neq[\empt]$, we know $S_0\FS(m)=\vald$.
Therefore $[S_0\FS(m)]=[z]=[\vald]$, which, combined with $y=\empt$
means that $\cxynv\cc\czwmv\nequiv\cbrk$ is a destruction pair, which is a contradiction.
\end{proof}

@@

\begin{mylem}\label{equiv_simple_same_commands}
If two simple sequences $A$ and $B$ are equivalent
and do not break all filesystems ($A\equiv B\nequiv\cbrk$),
then they must contain the same commands.
\end{mylem}
\begin{proof}
% BACK_TO_THE_FS
We use a proof by contradiction. Let $A$ and $B$ be simple sequences
such that $A\equiv B\nequiv\cbrk$,
and $\FS$ be a filesystem that they do not break.

@@

\begin{mylem}\label{simple_reorder_equiv}
For a simple sequence $S\nequiv\cbrk$,
$\orderset{S}$ is the set of all simple sequences equivalent to $S$, that is,
\[ \orderset{S} = [S]_\equiv \cap \setssmp. \]
\end{mylem}

@@

% TODO APPENDIX
\begin{proof}
First, we prove, by contradiction, that if a simple sequence $T\equiv S$, then $T\in\orderset{S}$.
We assume $T\equiv S$ and $T\not\in\orderset{S}$.
Then, from \cref{equiv_simple_same_commands} we know $T$ is a permutation of $S$,
and from \cref{order_is_only_possible} we know $T\equiv\cbrk$ which
is a contradiction as $T\equiv S\nequiv\cbrk$.

@@

Let us turn to the location of these commands in $T$,
and denote their indices in $T$ with $I^T_\indx$.
We know $I^T_k=I^T_0-1$ as $\czwmv$ precedes $\cxynv$.
This means there must exist a $\indx$ for which $I^T_{\indx+1}<I^T_\indx$.
We therefore know that
$\caaa{x_{\indx+1}}{y_{\indx+1}}{n_{\indx+1}} \cc \caaa{x_\indx}{y_\indx}{n_\indx}$
is a subsequence of $T$, and by assumption $\parentf{n_{\indx+1}}=n_\indx$.
By definition of $\ordersetsign$, therefore $y_{\indx+1}=\empt$ must be true,
which is a constradiction.

>>>>

\begin{mylem}\label{simple_distant_pairs}
For any simple sequence $\wrksx{S}$,
if $\cxynv\cc\czwmv$ is a subsequence of $S$ and $n=\parent(m)$ or $\parent(n)=m$,
then $\wrks{\cxynv\cc\czwmv}$.
\end{mylem}
\begin{proof}
Informally speaking, the proposition means that
the two commands must form a construction or destruction pair even
if they are not next to each other.
This is true because in simple sequences
there cannot be another command on $n$ or $m$,
so if the commands are incompatible, no command between them can change that.
In fact, this \namecref{simple_distant_pairs} is the general case
of \cref{ax_directchild_breaks,ax_directparent_breaks}.

% BACK_TO_THE_FS TODO
% TODO Review any proof depending on "nothing changes FS at n until...".
% TODO Move to rules
% TODO APPENDIX
Formally, we prove our proposition by contradiction,
and in this version of the proof we reach back to our filesystem model.
We assume $\wrks{S}$ and therefore there is a $\FS$ for which $S\aFS\neq\fsbroken$,
but $S$ contains commands on a node and its parent that do not form a construction or destruction pair.
Select two such commands, $\cxynv$ and $\czwmv$,
and split $S$ around them into three parts:
\[ S = S_0 \cc \cxynv \cc S_1 \cc \czwmv \cc S_2. \]
As $S$ is simple, there cannot be any commands on $n$ or $m$ in $S_1$,
and therefore $(S_0\cc\cxynv\cc S_1)\FS(n) = y$,
and $[(S_0\cc\cxynv\cc S_1)\FS(m)] = [S_0\FS(m)]$.
If $[z]\neq[S_0\FS(m)]$, then $S\aFS=\fsbroken$, which is a contradiction.

If $n=\parent(m)$, then 
we know $(S_0\cc\cxynv\cc S_1)\FS(n) = y$, and as $[z]\neq[w]$, either
$[(S_0\cc\cxynv\cc S_1)\FS(m)]=[z]\neq[\empt]$,
or $[(S_0\cc\cxynv\cc S_1\cc\czwmv)\FS(m)]=[w]\neq[\empt]$.
Therefore $y$ must be $\vald$, as otherwise the tree property would be violated
when applying $\czwmv$,
and $[x]\neq[\vald]$ as $\cxynv$ is not an assertion command.
As $[S_0\FS(n)]=[x]\neq[\vald]$, we know
$S_0\FS(m)=\empt$ as otherwise the tree property would be violated.
Therefore $[S_0\FS(m)]=[z]=[\empt]$, which, combined with $y=\vald$
means that $\wrksx{\cxynv\cc\czwmv}$ is a construction pair, which is a contradiction.

If $\parent(n)=m$, then 
as $(S_0\cc\cxynv\cc S_1)\FS(n) = y$, and either
$[(S_0\cc\cxynv\cc S_1)\FS(m)]=[z]\neq[\vald]$
or $[(S_0\cc\cxynv\cc S_1\cc\czwmv)\FS(m)]=[w]\neq[\vald]$,
$y$ must be $\empt$, and $[x]\neq[\empt]$.
As $[S_0\FS(n)]=[x]\neq[\empt]$, we know $S_0\FS(m)=\vald$.
Therefore $[S_0\FS(m)]=[z]=[\vald]$, which, combined with $y=\empt$
means that $\wrksx{\cxynv\cc\czwmv}$ is a destruction pair, which is a contradiction.
\end{proof}

@@

% TODO breaks --> not defined on

\begin{mylem}\label{equiv_simple_same_commands}
If two simple sequences $A$ and $B$ are equivalent
and do not break all filesystems ($A\equiv \wrksx{B}$),
then they must contain the same commands.
\end{mylem}
\begin{proof}
% BACK_TO_THE_FS
% TODO Rule?
We use a proof by contradiction. Let $A$ and $B$ be simple sequences
such that $A\equiv \wrksx{B}$,
and $\FS$ be a filesystem that they are defined on.

@@

\begin{mylem}\label{simple_reorder_equiv}
For a simple sequence $\wrksx{S}$,
$\orderset{S}$ is the set of all simple sequences equivalent to $S$, that is,
\[ \orderset{S} = [S]_\equiv \cap \setssmp. \]
\end{mylem}

@@

% TODO APPENDIX
\begin{proof}
First, we prove, by contradiction, that if a simple sequence $T\equiv S$, then $T\in\orderset{S}$.
We assume $T\equiv S$ and $T\not\in\orderset{S}$.
Then, from \cref{equiv_simple_same_commands} we know $T$ is a permutation of $S$,
and from \cref{order_is_only_possible} we know $T\equiv\cbrk$ which
is a contradiction as $T\equiv \wrksx{S}$.

@@

Let us turn to the location of these commands in $T$,
and denote their indices in $T$ with $I^T_\indx$.
We know $I^T_k=I^T_0-1$ as $\czwmv$ precedes $\cxynv$.
This means there must exist a $\indx$ for which $I^T_{\indx+1}<I^T_\indx$.
We therefore know that
$\caaa{x_{\indx+1}}{y_{\indx+1}}{n_{\indx+1}} \cc \caaa{x_\indx}{y_\indx}{n_\indx}$
is a subsequence of $T$, and by assumption $\parentf{n_{\indx+1}}=n_\indx$.
By definition of $\ordersetsign$, therefore $y_{\indx+1}=\empt$ must be true,
which is a contradiction.

### 2017-10-06 p2/works-inverse.tex
Remove unused symbols; clarify text

\begin{mydef}[Sets of sequences]
As we will frequently refer to sets of fequences,
we will use calligraphic letters (e.g. $\sqs{A}$, $\sqs{B}$, $\sqs{C}$ and $\sqs{S}$)
to denote such sets for brevity.
We will use $A\cc\sqs{B}$ to denote the set of sequences $\{A\cc B\whr B\in\sqs{B}\}$,
and $\Dom{\sqs{A}}$ to denote $\bigcap_{A\in\sqs{A}} \Dom{A}$.

We also write $\wrks{\sqs{A}}$ if $\varnothing \subset \Dom{\sqs{A}}$, that is,
if we know that there is at least one filesystem
all sequences in $\sqs{A}$ are defined on.
\end{mydef}

@@

\begin{mydef}[$\worksmsign$, $\worksmeqsign$]\label{def_works}
For two sets of sequences $\sqs{A}$ and $\sqs{B}$
we write $\worksm{\sqs{A}}{\sqs{B}}$ iff $\Dom{\sqs{A}} \subseteq \Dom{\sqs{B}}$;
that is, iff all sequences in $\sqs{B}$ are defined on (that is, do not break)
all filesystems sequences in $\sqs{A}$ are defined on.
We write $\worksmstr{\sqs{A}}{\sqs{B}}$ iff $\Dom{\sqs{A}} \subset \Dom{\sqs{B}}$.

>>>>

\begin{mydef}[Sets of sequences]
As we will frequently refer to sets of fequences,
we will use calligraphic letters (e.g. $\sqs{A}$, $\sqs{B}$, $\sqs{C}$ and $\sqs{S}$)
to denote such sets for brevity.
We will write $\Dom{\sqs{A}}$ for $\bigcap_{A\in\sqs{A}} \Dom{A}$,
and $\sqs{A}\cc\sqs{B}$ for $\{A\cc B\whr A\in\sqs{A}, B\in\sqs{B}\}$.
We will also use $A\cc\sqs{B}$ with the meaning $\{A\}\cc\sqs{B}$.

% This has only been defined for sequences, not sets.
Moreover, we write $\wrks{\sqs{A}}$ if $\varnothing \subset \Dom{\sqs{A}}$, that is,
if we know that there is at least one filesystem
all sequences in $\sqs{A}$ are defined on.
\end{mydef}

@@

\begin{mydef}[$\worksmeqsign$]\label{def_works}
For two sets of sequences $\sqs{A}$ and $\sqs{B}$
we write $\worksm{\sqs{A}}{\sqs{B}}$ iff $\Dom{\sqs{A}} \subseteq \Dom{\sqs{B}}$;
that is, iff all sequences in $\sqs{B}$ are defined on (do not break)
all filesystems on which sequences in $\sqs{A}$ are defined.
